---
title: "Single fit and overlay"
author: "h qin"
date: "June 9, July 6-11, 2017"
output:
  html_document: default
  pdf_document: default
---

20171002 Add AIC 
The log-likelihood values calculated by FlexSurv and my own routine return values with 10x differences. This is an old problem. 
see http://hongqinlab.blogspot.com/2014/07/compare-binomial-gompertz-and-weibull.html 
It seems FlexSurv used the loglikehood value from optim(). When I switch to optim returned $value, the results are comparable. 

20170424. Major bug found. file names and strains names are inconsistent. 

20160711-> Fitting RLS dataset merged by strains. 

20160706-0711. Finished batch fitting of all individual RLS data sets. The fitting results showed that 'n' are often in the range of [5,7], though it is very noisy. 


```{r}
rm(list=ls())
#setwd("~/github/0.network.aging.prj.bmc/0a.rls.fitting")
#setwd("~/github/bmc_netwk_aging_manuscript/R1/0.nat.rls.fitting")
library('flexsurv')
source("../lifespan.r")
```

```{r}
##### log likelihood function, R-fixed 2-parameter binomial mortality rate model
# http://hongqinlab.blogspot.com/2013/12/binomial-mortailty-model.html
# m = R ( 1 + t/t0)^(n-1)
# s = exp( (R t0/n)*(1 - (1+t/t0)^n ) )  

llh.binomialMortality.single.run.NoR0 <- function( Rt0n, lifespan, debug=0 ) {
  I = R0;
  t0 = Rt0n[2]; n=Rt0n[3];
  my.data = lifespan[!is.na(lifespan)];
  log_s = (I * t0 /n )*(1 - (1 + my.data/t0)^n);
  log_m = log(I) +  (n-1) * log(1 + my.data/t0 ); 
  my.lh = sum(log_s)  + sum(log_m);
  if(debug) { print (c(I, Rt0n )); } #trace the convergence
  ret = - my.lh # because optim seems to maximize 
}


llh.binomialMortality.single.run.FixedN <- function( Rt0n, lifespan, debug=0 ) {
  I=Rt0n[1]; t0 = Rt0n[2]; #n=Rt0n[3]; n defined outside
  my.data = lifespan[!is.na(lifespan)];
  log_s = (I * t0 /n )*(1 - (1 + my.data/t0)^n);
  log_m = log(I) +  (n-1) * log(1 + my.data/t0 ); 
  my.lh = sum(log_s)  + sum(log_m);
  if(debug) { print (c(I, Rt0n )); } #trace the convergence
  ret = - my.lh # because optim seems to maximize 
}


```

# Parse  strains from files

```{r}
set.seed(20170101)
```

Take files from natural isolates
```{r}
my.strains=c("101S", "M1-2","M13","M14","M2-8","M22","M32","M34","M5","M8","RM112N","S288c","SGU57", "YPS128","YPS163")

report = data.frame(cbind(my.strains))
report$samplesize = NA; report$R=NA; report$t0=NA; report$n=NA; report$G=NA; report$longfilename=NA; 
```

# Now, fit all RLS data sets by strains

```{r, message=FALSE }
for( i in 1:length(report[,1])){
#for( i in 3:4){
  filename = paste("rls/", my.strains[i], ".csv", sep='' ); 
  report$longfilename[i] = filename; 
  tb = read.csv( filename, header =F)

  report$samplesize[i] = length(tb[,1])

  GompFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'gompertz')
  WeibFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'weibull')

  report$avgLS[i] =  mean(tb[,1])
  report$stdLS[i] =  sd(tb[,1])
  report$CV[i] = report$stdLS[i] / report$avgLS[i]

  report$GompGFlex[i] = GompFlex$res[1,1]
  report$GompRFlex[i] = GompFlex$res[2,1]
  report$GompLogLikFlex[i] = round(GompFlex$loglik, 1)
  report$GompAICFlex[i] = round(GompFlex$AIC)

  report$WeibShapeFlex[i] = WeibFlex$res[1,1]
  report$WeibRateFlex[i] = WeibFlex$res[2,1]
  report$WeibLogLikFlex[i] = round(WeibFlex$loglik, 1)  
  report$WeibAICFlex[i] = round(WeibFlex$AIC)

  #set initial values
  Rhat = report$GompRFlex[i]; # 'i' was missing. a bug costed HQ a whole afternoon.
  Ghat = report$GompGFlex[i];
  nhat = 8;  
  t0= (nhat-1)/Ghat;
  
  if( is.na( match(my.strains[i], c('M8' )) ) ) {
    fitBinom = optim ( c(Rhat, t0, nhat),  llh.binomialMortality.single.run, 
                       lifespan=tb[,1], 
                       #method='SANN') #SANN needs control  
                       method="L-BFGS-B", 
                       lower=c(1E-4, 1, 2), upper=c(0.1,200,100) );  
    report[i, c("R", "t0", "n")] = fitBinom$par[1:3]
    report$BinomialAIC[i] = 2*3 + 2*fitBinom$value
  } else { #M8, fix n
    print(my.strains[i])
    #llh.binomialMortality.single.run.FixedN <- function( Rt0n, lifespan, debug=0 ) {
    n=11.5;
    fitBinom = optim ( c(Rhat, t0, nhat),  llh.binomialMortality.single.run.FixedN, 
                       lifespan=tb[,1], 
                       #method='SANN') #SANN needs control  
                       method="L-BFGS-B", 
                       lower=c(1E-5, 1, 2), upper=c(0.1,200,100) );  
    report[i, c("R", "t0", "n")] = c(fitBinom$par[1:2], n)
    report$BinomialAIC[i] = 2*3 + 2*fitBinom$value
  }
  report$G[i] = (report$n[i] - 1)/report$t0[i]

  #AIC 
  # llh.binomialMortality.single.run <- function( Rt0n, lifespan, debug=0 )
  #report$BinomialAICqin[i] = 2*3 + 2*llh.binomialMortality.single.run( report[i, c("R", "t0", "n")], lifespan=tb[,1] )

  #llh.G.single.run <- function( IG, lifespan ) {
  #report$GompAICqin[i] = 2*2 + 2* llh.G.single.run(report[i,c("GompRFlex", "GompGFlex")], lifespan=tb[,1])
    # dgompertz <- function(x, shape, rate=1, log=FALSE) 
  #report$GompAICdgompergtz[i] = 2*2 - 2* sum(dgompertz(tb[,1],shape=report$GompGFlex[i], rate=report$GompRFlex, log=TRUE));
  #report$GompAICdgompergtz2[i] = 2*2 - 2* sum(dgompertz(tb[,1],shape=report$GompRFlex[i], rate=report$GompGFlex, log=TRUE));

}
```

Show the results
```{r}
#report[ grep("tBY", report$strains), ]
report
summary(report[, c("avgLS", "t0", "n")])
```

#Overlay the binomial aging model with observation. 
see  http://hongqinlab.blogspot.com/2013/12/binomial-mortailty-model.html
 m = R ( 1 + t/t0)^(n-1)
 s = exp( (R t0/n)*(1 - (1+t/t0)^n ) )  

```{r, message=FALSE }
for( i in 1:length(my.strains)){
#for( i in 3:4){
  filename = paste("rls/", my.strains[i], ".csv", sep='' ); 
  tb = read.csv( filename, header =F)
  
  ret = calculate.s(tb[,1])
  plot( ret$s ~ ret$t, main=my.strains[i]);
  print (report[i,  ]);

  #overlay binomial aging viability
  print (report[i, c("R", "t0", "n", "G")] );
  t = seq(0,max(ret$t*1.1), 0.1);
  # s = exp( (R t0/n)*(1 - (1+t/t0)^n ) )  
  s = exp( (report$R[i]* report$t0[i]/report$n[i])*(1 - (1+t/report$t0[i])^report$n[i] ) ) ; 
  lines(s ~ t, col='red')
  
  #overlay gompertz  viability
  s.g =  G.s( c(report$GompRFlex[i], report$GompGFlex[i], 0), t )
  lines(s.g ~ t, col='blue')
 }
 


```

#show the report
```{r}
report
```

Calculate lambda based on t0 = (1-p)/(p Lambda)
So, 1/lambda = t0 * p / (1-p)

```{r}
p = 0.7
report$One.over.lambdaP07 = report$t0 * p/ (1-p)
report[, c("t0", "One.over.lambdaP07")]
summary( report[, c("t0", "One.over.lambdaP07")])

```

Histogram of 1/lambda for p=0.7
```{r}
hist(report$One.over.lambdaP07, br=20, xlim=c(60,190), xlab=expression(paste("1/",lambda) ), main=expression(paste("Histogram of 1/",lambda, " when p=0.7")))
```

```{r}
p = 0.9
report$One.over.lambdaP09 = report$t0 * p/ (1-p)
report[, c("t0", "One.over.lambdaP09")]
summary( report[, c("t0", "One.over.lambdaP09")] )
```

Histogram 1/lambda, p=0.9
```{r}
hist(report$One.over.lambdaP09, br=20, xlim=c(200,800), xlab=expression(paste("1/",lambda) ), main=expression(paste("Histogram of 1/",lambda, " when p=0.9")))
```




# Explorative analysis 

```{r}
# my.strains=c("101S", "BY4743","M1-2","M13","M14","M2-8","M22","M32","M34","M5","M8","RM112N","S288c","SGU57","SK1","W303","YPS128","YPS163")

#my.strains=c("101S", "M1-2","M13","M14","M2-8","M22","M32","M34","M5","M8","RM112N","S288c","SGU57", "YPS128","YPS163")
report2 = report
summary(lm(log10(report2$GompRFlex) ~ report2$GompGFlex))
summary(lm(log10(report2$R) ~ report2$GompGFlex))
summary(lm(log10(report2$R) ~ report2$G)) #G from t0 and n 
summary(lm(report2$GompGFlex ~ report2$G)) #good agreement bwteen GFlex and G from binomial modeling

summary(lm(log10(report2$R) ~ report2$t0)) #G from t0 and n 
summary(lm(log10(report2$R) ~ report2$n)) #G from t0 and n 

```

# Strehler-Mildvan correlation in natural isolates
```{r}
plot( log10(report2$R) ~ report2$G, pch=19, col='red', xlab='G', ylab='log10(R)', 
      xlim=c(0.07, 0.22), ylim=c(-2.9, -2.1),
      main="Strehler-Mildvan correlation in yeast wild isolates")
m = lm(log10(report2$R) ~ report2$G)
abline (m, col='black', lty=2);
text(0.18, -2.2, "Rsquared=0.39 p=0.01")

my.x = report2$G +0.0013 ; my.y = log10(report2$R) ; 
names(my.x) = report2$my.strains; names(my.y) = report2$my.strains; 
my.y[c("YPS163", "M1-2" )]=c(-2.8, -2.63)
text(my.x,my.y, report2$my.strains, adj=c(0,0) )


summary(lm(log10(report2$R) ~ report2$G)) #G from t0 and n 

summary(lm(log10(report2$GompRFlex) ~ report2$GompGFlex)) #G from t0 and n 

```


# plot log10(R) ~ t0

```{r}
plot( log10(report2$R) ~ report2$t0, col='red',  pch=19, xlab='t0', ylab='log10(RFlex)', ylim=c(-3,-2))
my.x = report2$t0; my.y = log10(report2$R) + 0.04; 
names(my.x) = report2$my.strains; names(my.y) = report2$my.strains; 
my.y[c("M32","YPS163", "YPS128", "M1-2" )]=c(-2.75, -2.8, -2.75, -2.63)
text(my.x,my.y, report2$my.strains )
m = lm(  log10(report2$R) ~ report2$t0  )
abline (m, col='black', lty=2);
text( 60, -2.8, "R^2=0.44, p=0.007")
summary(m)
```

# plot log10(GompRFlex) ~ t0

```{r}
plot( log10(report2$GompRFlex) ~ report2$t0, col='red',  pch=19, xlab='t0', ylab='log10(Rflex)', ylim=c(-3.5,-2))
my.x = report2$t0; my.y = log10(report2$GompRFlex) + 0.04; 
names(my.x) = report2$my.strains; names(my.y) = report2$my.strains; 
my.y[c("M32","YPS163", "YPS128", "M1-2" )]=c(-2.75, -2.8, -2.75, -2.63)
text(my.x,my.y, report2$my.strains )
m = lm(  log10(report2$GompRFlex) ~ report2$t0  )
abline (m, col='black', lty=2);
text( 60, -2.8, "R^2=0.36, p=0.019")
summary(m)
```


# Could t0 be the cause of logR-G correlation? 
Partial regression on G ~  log(R)  + t0. 

```{r}
 summary(lm( log10(report2$GompRFlex) ~ report2$t0 + report2$GompGFlex  )) #poor
 summary(lm( log10(report2$GompGFlex) ~ report2$t0 + report2$GompRFlex  ))#good
#summary(lm( report2$G ~ log10(report2$R) + report2$t0  ))
#summary(lm( report2$G ~ report2$t0 + log10(report2$R)   ))
summary(lm( report2$GompGFlex ~ report2$t0 + log10(report2$R)   )) 
summary(lm( report2$GompGFlex ~ report2$t0 + log10(report2$GompRFlex)   )) 

```


# Mediation test on Gflex <--t0 <-- RFlex
Hong thinks this gives positive result. 
```{r}
library(mediation)
set.seed(20170801)
report2$log10GompRFlex = log10(report2$GompRFlex)
med.fit = lm(t0 ~ log10GompRFlex, data=report2)  
summary(med.fit)
out.fit = lm(GompGFlex ~ t0 + log10GompRFlex, data=report2)  
summary(out.fit)
med.out <- mediate(med.fit, out.fit, treat = "log10GompRFlex", mediator = "t0", robustSE = TRUE, sims = 100)
summary(med.out)

```

# Mediation test 2 on Rflex <--t0 <-- GFlex
Hong thinks this is negative result. 
```{r}
med.fit = lm(t0 ~ GompGFlex, data=report2)  
summary(med.fit)
out.fit = lm(log10GompRFlex ~ t0 + GompGFlex, data=report2)  
summary(out.fit)
med.out <- mediate(med.fit, out.fit, treat = "GompGFlex", mediator = "t0", robustSE = TRUE, sims = 100)
summary(med.out)


```

# Plot initial virtula life ~ n (average interactions)

```{r}
plot( report2$t0 ~ report2$n, col='red', xlim=c(5.5, 9), ylim=c(25, 80), pch=19, xlab='n', ylab='t0')
my.x = report2$n + 0.1; my.y = report2$t0 + 2; 
names(my.x) = report2$my.strains; names(my.y) = report2$my.strains; 
my.y[c("M14","SGU57", "M32","M13" )]=c(52, 52, 33, 38)
text(my.x,my.y, report2$my.strains )
text(6, 70, "p=0.60")
summary(lm(log10(report2$t0) ~ report2$n)) #G from t0 and n 

```

Gompert versus Weibull?
AIC: smaller is better (for information loss)
```{r}
report$BestModel = ifelse(report$GompAICFlex < report$WeibAICFlex, "Gomp", "Weib")
report$BestModel = ifelse(abs(report$GompAICFlex - report$WeibAICFlex)<2, "<2", report$BestModel)
```


CV ~ Gomp and Weibull? How does noises influence likelihood of Gompertz and Weibull fitting? 
```{r}
summary(lm(report$CV ~ report$BestModel ))
summary(lm(report$CV ~ report$WeibLogLikFlex ))
summary(lm(report$CV ~ report$GompLogLikFlex  ))
summary(lm(report$CV ~ (report$GompLogLikFlex - report$WeibLogLikFlex)))
plot( report$GompLogLikFlex ~ report$CV, col="red", pch=3, xlim=c(0, 0.8), ylim=c(-220, -80))
points( report$CV, report$WeibLogLikFlex, col="blue", pch=4)
m1 = lm( report$GompLogLikFlex ~ report$CV)
m2 = lm( report$WeibLogLikFlex ~ report$CV)
abline( m1, col="red", lty=2)
abline( m2, col='blue', lty=1)
text(0.6, -210, "nearly the same!?")
```

The QIN-RLS data suggested that noisy system signal perfer Gompertz model, based on GG01 theory. 
Notice that CV measures distrubition of system signals and are different from white noises (residues)

```{r}
report$DeltaGompWeiLLH = report$GompLogLikFlex - report$WeibLogLikFlex
plot( report$DeltaGompWeiLLH ~ report$CV )
m3 = lm( report$DeltaGompWeiLLH ~ report$CV)
abline(m3, col='red')
```


