---
title: "0.fit_qinlabrls.Rmd"
author: "h qin"
date: "June 9, July 6, 2016"
output: html_document
---

```{r}
rm(list=ls())
setwd("~/github/0.network.aging.prj.bmc/0a.rls.fitting")
library('flexsurv')
source("lifespan.r")
```

# First, explore the fitting outcomes of 'flexsurv'. 

```{r}
files = list.files(path="../qinlab_rls/", pattern="rls.tab")
report = data.frame(cbind(files))
report$R=NA; report$t0=NA; report$n=NA;

i=2
  tb = read.table( paste("../qinlab_rls/",files[i],sep=''), sep="\t")
  GompFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'gompertz')
  WeibFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'weibull')
str(GompFlex)
GompFlex$res
GompFlex$res.t
GompFlex$opt$hessian

str(WeibFlex)

```

# Now, fit all RLS data sets

```{r, message=FALSE }
for( i in 1:length(report[,1])){
#for( i in 3:4){
  tb = read.table( paste("../qinlab_rls/",files[i],sep=''), sep="\t")
  GompFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'gompertz')
  WeibFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'weibull')

  report$avgLS[i] =  mean(tb[,1])
  report$stdLS[i] =  sd(tb[,1])
  report$CV[i] = report$stdLS[i] / report$avgLS[i]

  report$GompGFlex[i] = GompFlex$res[1,1]
  report$GompRFlex[i] = GompFlex$res[2,1]
  report$GompLogLikFlex[i] = round(GompFlex$loglik, 1)
  report$GompAICFlex[i] = round(GompFlex$AIC)

  report$WeibShapeFlex[i] = WeibFlex$res[1,1]
  report$WeibRateFlex[i] = WeibFlex$res[2,1]
  report$WeibLogLikFlex[i] = round(WeibFlex$loglik, 1)  
  report$WeibAICFlex[i] = round(WeibFlex$AIC)

  #set initial values
  Rhat = report$GompRFlex[i]; # 'i' was missing. a bug costed HQ a whole afternoon.
  Ghat = report$GompGFlex[i];
  nhat = 6;  
  t0= (nhat-1)/Ghat;
  fitBinom = optim ( c(Rhat, t0, nhat),  llh.binomialMortality.single.run, 
                     lifespan=tb[,1], 
                     #method='SANN') #SANN needs control  
                     method="L-BFGS-B", 
                     lower=c(1E-10, 1, 1), upper=c(1,200,20) );  
  
  report[i, c("R", "t0", "n")] = fitBinom$par[1:3]
}

```

Show the results
```{r}
report[ grep("BY", report$files), ]
#report[,c("files", "BestModel","CV", "GompAICFlex", "WeibAICFlex", "GompLogLikFlex", "WeibLogLikFlex")]
#report[,c("files", "BestModel", "GompAICFlex", "WeibAICFlex", "GompLogLikFlex", "WeibLogLikFlex")]
write.csv(report, file = 'sandbox/_report_qinlab_rls.csv', row.names = FALSE)
```


Gompert versus Weibull?
AIC: smaller is better (for information loss)
```{r}
report$BestModel = ifelse(report$GompAICFlex < report$WeibAICFlex, "Gomp", "Weib")
report$BestModel = ifelse(abs(report$GompAICFlex - report$WeibAICFlex)<2, "<2", report$BestModel)
```


CV ~ Gomp and Weibull? How does noises influence likelihood of Gompertz and Weibull fitting? 
```{r}
summary(lm(report$CV ~ report$BestModel ))
summary(lm(report$CV ~ report$WeibLogLikFlex ))
summary(lm(report$CV ~ report$GompLogLikFlex  ))
summary(lm(report$CV ~ (report$GompLogLikFlex - report$WeibLogLikFlex)))
plot( report$GompLogLikFlex ~ report$CV, col="red", pch=3, xlim=c(0, 0.8), ylim=c(-220, -80))
points( report$CV, report$WeibLogLikFlex, col="blue", pch=4)
m1 = lm( report$GompLogLikFlex ~ report$CV)
m2 = lm( report$WeibLogLikFlex ~ report$CV)
abline( m1, col="red", lty=2)
abline( m2, col='blue', lty=1)
text(0.6, -210, "nearly the same!?")
```

The QIN-RLS data suggested that noisy system signal perfer Gompertz model, based on GG01 theory. 
Notice that CV measures distrubition of system signals and are different from white noises (residues)

```{r}
report$DeltaGompWeiLLH = report$GompLogLikFlex - report$WeibLogLikFlex
plot( report$DeltaGompWeiLLH ~ report$CV )
m3 = lm( report$DeltaGompWeiLLH ~ report$CV)
abline(m3, col='red')
```


TODO: Calculate the white noises (fitting errors) using the fitting residues for Gompertz and Weibull models. 

```{r}
# report$residues ?? 

```




