---
title: "Mediation test on natural isolates. short version"
author: "h qin"
date: "June 13, 2017"
output:
  html_document: default
  pdf_document: default
---


```{r}
rm(list=ls())
set.seed(20170101)
#setwd("~/github/0.network.aging.prj.bmc/0a.rls.fitting")
setwd("~/github/bmc_netwk_aging_manuscript/R1/0.nat.rls.fitting")
library('flexsurv')
source("../lifespan.r")
```

# Parse  strains from files

```{r}
files = list.files(path="../qinlab_rls/", pattern="rls.tab")
tmp1 = gsub("\\d{6}.", "", files)
redundant_strains = gsub(".rls.tab", "", tmp1)
strains = sort( unique( redundant_strains ))
strains
```

Take files from natural isolates
```{r}
my.strains=c("101S", "M1-2","M13","M14","M2-8","M22","M32","M34","M5","M8","RM112N","S288c","SGU57", "YPS128","YPS163")
files2=c();
for( i in 1:length(my.strains)){
 files2 = c( files2, files[grep(my.strains[i], files)]);
}

report = data.frame(cbind(my.strains))
report$samplesize = NA; report$R=NA; report$t0=NA; report$n=NA; report$G=NA; report$longfilename=NA; 

files = files2; 
strains = my.strains; 
```

# Now, fit all RLS data sets by strains

```{r, message=FALSE }
for( i in 1:length(report[,1])){
#for( i in 3:4){
  my.files = files[grep(strains[i], files)]
  report$longfilename[i] = paste(my.files, collapse = "::");
  tb = read.table( paste("../qinlab_rls/",my.files[1],sep=''), sep="\t")
  if( length(my.files)> 1){
    for( fi in 2:length(my.files)) {
      tmp.tb = read.table( paste("../qinlab_rls/",my.files[fi],sep=''), sep="\t")
      tb = rbind( tb, tmp.tb)
    }
  }
  report$samplesize[i] = length(tb[,1])

  GompFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'gompertz')
  WeibFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'weibull')

  report$avgLS[i] =  mean(tb[,1])
  report$stdLS[i] =  sd(tb[,1])
  report$CV[i] = report$stdLS[i] / report$avgLS[i]

  report$GompGFlex[i] = GompFlex$res[1,1]
  report$GompRFlex[i] = GompFlex$res[2,1]
  report$GompLogLikFlex[i] = round(GompFlex$loglik, 1)
  report$GompAICFlex[i] = round(GompFlex$AIC)

  report$WeibShapeFlex[i] = WeibFlex$res[1,1]
  report$WeibRateFlex[i] = WeibFlex$res[2,1]
  report$WeibLogLikFlex[i] = round(WeibFlex$loglik, 1)  
  report$WeibAICFlex[i] = round(WeibFlex$AIC)

  #set initial values
  Rhat = report$GompRFlex[i]; # 'i' was missing. a bug costed HQ a whole afternoon.
  Ghat = report$GompGFlex[i];
  nhat = 6;  
  t0= (nhat-1)/Ghat;
  fitBinom = optim ( c(Rhat, t0, nhat),  llh.binomialMortality.single.run, 
                     lifespan=tb[,1], 
                     #method='SANN') #SANN needs control  
                     method="L-BFGS-B", 
                     lower=c(1E-10, 1, 4), upper=c(1,200,20) );  
  report[i, c("R", "t0", "n")] = fitBinom$par[1:3]
  report$G[i] = (report$n[i] - 1)/report$t0[i]
}
report2 = report; 
```


# Mediation test on Gflex <--t0 <-- RFlex
# Hong thinks the results indicate the t0 is the mediator from Flex to GFlex, but not sure.  

```{r}
library(mediation)
set.seed(20170801)
report2$log10GompRFlex = log10(report2$GompRFlex)
med.fit = lm(t0 ~ log10GompRFlex, data=report2)  
summary(med.fit)
out.fit = lm(GompGFlex ~ t0 + log10GompRFlex, data=report2)  
summary(out.fit)
med.out <- mediate(med.fit, out.fit, treat = "log10GompRFlex", mediator = "t0", robustSE = TRUE, sims = 100)
summary(med.out)

```

# Mediation test 2 on Rflex <--t0 <-- GFlex
#Hong thinks this is negative result, which means t0 works only in one direction. 

```{r}
med.fit = lm(t0 ~ GompGFlex, data=report2)  
summary(med.fit)
out.fit = lm(log10GompRFlex ~ t0 + GompGFlex, data=report2)  
summary(out.fit)
med.out <- mediate(med.fit, out.fit, treat = "GompGFlex", mediator = "t0", robustSE = TRUE, sims = 100)
summary(med.out)


```


#overlay binomial aging model cuvrve with RLS data

