---
title: "Fit Qin-lab natural isolates using bootstraps"
author: "h qin"
date: "2017 April 24 - June 11"
output:
  html_document: default
  pdf_document: default
---


```{r}
rm(list=ls())
library('flexsurv')
source("lifespan.r")
```

Set seed for consistency for publication purpose.  
```{r}
set.seed(20170101)  #remove to test for different bootstraps
```

# parse the strains from files

```{r}
RUNS = 100; #bootstrap runs
strains=c("101S", "M1-2","M13","M14","M2-8","M22","M32","M34","M5","M8","RM112N","S288c","SGU57", "YPS128","YPS163")

report = data.frame(cbind(strains))
report$samplesize = NA; report$R=NA; report$t0=NA; report$n=NA; report$G=NA; report$longfilename=NA; 

```


# Now, fit all RLS data sets by strains, and output each bootstrap to a single csv file. 

```{r fit_all, message=FALSE }
for( BootstrapCount in 1:RUNS ) {

for( i in 1:length(report[,1])){
#for( i in 3:4){
  filename = paste("rls/", strains[i], ".csv", sep='' ); 
  report$longfilename[i] = filename; 
  tb = read.csv( filename, header= F)
  report$samplesize[i] = length(tb[,1])

  #boostrap 
  tb[,1] = sample(tb[,1], replace=TRUE); # BOOTSTRAP HERE
  
  GompFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'gompertz')
  WeibFlex = flexsurvreg(formula = Surv(tb[,1]) ~ 1, dist = 'weibull')

  report$avgLS[i] =  mean(tb[,1])
  report$stdLS[i] =  sd(tb[,1])
  report$CV[i] = report$stdLS[i] / report$avgLS[i]

  report$GompGFlex[i] = GompFlex$res[1,1]
  report$GompRFlex[i] = GompFlex$res[2,1]
  report$GompLogLikFlex[i] = round(GompFlex$loglik, 1)
  report$GompAICFlex[i] = round(GompFlex$AIC)

  report$WeibShapeFlex[i] = WeibFlex$res[1,1]
  report$WeibRateFlex[i] = WeibFlex$res[2,1]
  report$WeibLogLikFlex[i] = round(WeibFlex$loglik, 1)  
  report$WeibAICFlex[i] = round(WeibFlex$AIC)

  #set initial values
  Rhat = report$GompRFlex[i]; # 'i' was missing. a bug costed HQ a whole afternoon.
  Ghat = report$GompGFlex[i];
  nhat = 6;  
  t0= (nhat-1)/Ghat;
  fitBinom = optim ( c(Rhat, t0, nhat),  llh.binomialMortality.single.run, 
                     lifespan=tb[,1], 
                     #method='SANN') #SANN needs control  
                     method="L-BFGS-B", 
                     lower=c(1E-10, 1, 4), upper=c(1,200,20) );  
  report[i, c("R", "t0", "n")] = fitBinom$par[1:3]
  report$G[i] = (report$n[i] - 1)/report$t0[i]
  report$BinomialAIC[i] = 2*3 + 2*fitBinom$value #20171001 AIC for 3-parameter binomial aging model
  
}#i loop
  
report$BestModel = 'NA';
report$BestModel = ifelse((report$GompAICFlex < report$WeibAICFlex)&(report$GompAICFlex < report$BinomialAIC), "Gomp", report$BestModel)
report$BestModel = ifelse((report$BinomialAIC < report$WeibAICFlex)&(report$BinomialAIC < report$GompAICFlex), "Binom", report$BestModel)
report$BestModel = ifelse((report$WeibAICFlex < report$GompAICFlex)&(report$WeibAICFlex < report$BinomialAIC), "Weib", report$BestModel)
report$BestModel = ifelse( (abs(report$BinomialAIC - report$GompAICFlex)<2 ) | (abs(report$BinomialAIC-report$WeibAICFlex)<2 ), "bin<2", report$BestModel)

outname = paste("bootstrap/", BootstrapCount, ".csv", sep='');
write.csv(report, file = outname, row.names = FALSE);
}#end of boostrap loop

```

#summarize boostrap results. 
Pick row-col value from every file into a buffer, then mean and stddev. 

```{r}
BootstrapMean = report; 
BootstrapStd = report;
rownames = names(report);
for( col in 2:length(report[1,])) {
  for ( row in 1:length(report[,1])) {
    buffer = c();
    for( BootstrapCount in 1:RUNS ) { #merge all cells into buffer
      filename = paste("bootstrap/", BootstrapCount, ".csv", sep='');
      tb = read.csv(filename)
      if( rownames[col] == "BestModel" ) {  #I ignored the longfile name columns for the summary output
        buffer = as.character( c(buffer, as.character(tb[row,col])) );
      } else {
        buffer = c(buffer, tb[row, col]); 
      }
    }
    #get mean or summary info from buffer
    if( rownames[col] == "BestModel" ) {
      tmp = table( buffer );
      BootstrapMean[row,col] = paste(names(tmp), tmp, sep="=", collapse = ":");
    } else {
     BootstrapMean[row,col] = mean(buffer);
     BootstrapStd[row,col] = sqrt(var(buffer));
    }
  }  
}

```

Means 
```{r}
BootstrapMean
```


```{r}
summary(BootstrapMean$n)
```

StdDev
```{r}
BootstrapStd
```

Merge the two tables
```{r}
BootstrapMean$Rstd = BootstrapStd$R
BootstrapMean$t0std = BootstrapStd$t0
BootstrapMean$nstd = BootstrapStd$n
BootstrapMean$Gstd = BootstrapStd$G
BootstrapMean$avgLSstd = BootstrapStd$avgLS
names(BootstrapMean)
```

Reorganized the columns
```{r}
BootstrapMean = BootstrapMean[, c( "strains",  "samplesize",  "R", "Rstd",  "t0",  "t0std",  "n",  "nstd",   "G",  
                                   "Gstd", "avgLS",  "avgLSstd", "stdLS",   "BestModel",  
                                   "longfilename",   "CV",  "GompGFlex",  "GompRFlex",  "GompLogLikFlex", 
 "WeibShapeFlex",  "WeibRateFlex",  "WeibLogLikFlex",  "WeibAICFlex", "GompAICFlex", "BinomialAIC"  )];
BootstrapMean
```

Merge mean and std for publication
```{r}
BootstrapMeanPublishing = data.frame( BootstrapMean[, c("strains")] )
BootstrapMeanPublishing$RwithStd = as.vector( paste(round(BootstrapMean$R, 4), round(BootstrapMean$Rstd,4), sep=" +/- ") );
BootstrapMeanPublishing$t0withStd = as.vector( paste(round(BootstrapMean$t0, 1), round(BootstrapMean$t0std,1), sep=" +/- ") );
BootstrapMeanPublishing$nwithStd = as.vector( paste(round(BootstrapMean$n, 1), round(BootstrapMean$nstd,3), sep=" +/- ") );
BootstrapMeanPublishing$GwithStd = as.vector( paste(round(BootstrapMean$G, 2), round(BootstrapMean$Gstd,3), sep=" +/- ") );
BootstrapMeanPublishing$avgLSwithStd = as.vector( paste(round(BootstrapMean$avgLS, 2), round(BootstrapMean$avgLSstd,3), sep=" +/- ") );

BootstrapMeanPublishing
```

output
```{r}
write.csv(BootstrapMean, file="sandbox/Bootstrap_summary.csv", row.names = FALSE)
write.csv(BootstrapMeanPublishing, file="sandbox/Bootstrap_summary_for_publication.csv", row.names = FALSE)

```

Calculate lambda based on t0 = (1-p)/(p Lambda)
So, 1/lambda = t0 * p / (1-p)

```{r}
p = 0.7
BootstrapMean$One.over.lambdaP07 = BootstrapMean$t0 * p/ (1-p)
BootstrapMean[, c("t0", "One.over.lambdaP07")]
```

Histogram of 1/lambda for p=0.7
```{r}
hist(BootstrapMean$One.over.lambdaP07, br=20, xlim=c(60,190), xlab=expression(paste("1/",lambda) ), main=expression(paste("Histogram of 1/",lambda, " when p=0.7")))
```

```{r}
p = 0.9
BootstrapMean$One.over.lambdaP09 = BootstrapMean$t0 * p/ (1-p)
BootstrapMean[, c("t0", "One.over.lambdaP09")]
```

